Application of Turchin's method of statistical regularization

### 2018-03-08 AYSS-2017/article/ayss2017-Zeleniy.tex
Последние правки перед отправкой

Consider the usual situation: detector measures a signal ($\varphi(x)$) and produces observed data ($f(y)$) via convolution with its own apparatus function ($K(x,y)$). The resulting observed data $f(y)$ contains a random noise factor both from initial statistical uncertainty of $\varphi(x)$ and additional noise from measurement procedure. In order to reconstruct initial signal, one needs to solve Fredholm equation:
\begin{equation}
\label{eq-01}
    f(y) = \int dx~ K(x,y)\varphi(x),
\end{equation}
but this equation is ill-possed: a small error in the measurement of $f(y)$ leads to big instability of $\varphi(x)$.

@@

Solving such ill-possed problems requires additional operation called regularization. Regularization is a process of introducing additional information for transition from ill-possed problem to well-possed problem. The idea of statistical regularization~\cite{Turchin, Turovceva} is to look on the problem from the point of view of Bayessian statistics approach: unknown statistical value $\varphi(x)$ could be reconstructed using  measured value $f(y)$, the model ($K(x,y)$) and some prior information about $\varphi(x)$ behavior. Main features of statistical regularization:
\begin{itemize}
    \item Based on Bayesian approach and decision theory (choice theory).
    \item It's able to accommodate various forms of prior information, such as
      smoothness, constraints on boundary conditions, non-negativity, etc.
    \item It allows estimation of errors for reconstructed function $\varphi$
    \item Gives clear data-driven procedure for choice of regularization
      strength.
\end{itemize}

@@

Since original work is very hard to find and is not accessible in languages other then Russia, we will present here the short description of the method.

>>>>

Consider the usual situation: detector measures a signal ($\varphi(x)$) and produces observed data ($f(y)$) via convolution with its own apparatus function ($K(x,y)$). The resulting observed data $f(y)$ contains a random noise factor both from initial statistical uncertainty of $\varphi(x)$ and additional noise from measurement procedure. In order to reconstruct initial signal, one needs to solve Fredholm equation of 1-st kind:
\begin{equation}
\label{eq-01}
    f(y) = \int dx~ K(x,y)\varphi(x),
\end{equation}
but this equation is ill-possed: a small error in the measurement of $f(y)$ leads to big instability of $\varphi(x)$.

@@

Solving such ill-possed problems requires additional operation called regularization. It means we need to introduce additional information to problem well-posed one. The idea of statistical regularization~\cite{Turchin, Turovceva} is to look on the problem from the point of view of Bayessian statistics approach: unknown statistical value $\varphi(x)$ could be reconstructed using  measured value $f(y)$, the model ($K(x,y)$) and some prior information about $\varphi(x)$ behavior. Main features of statistical regularization:
\begin{itemize}
    \item Based on Bayesian approach and decision theory (choice theory).
    \item It's able to accommodate various forms of prior information, such as
      smoothness, constraints on boundary conditions, non-negativity, etc.
    \item It allows estimation of errors for reconstructed function $\varphi$
    \item Gives clear data-driven procedure for choice of regularization
      strength.
\end{itemize}

@@

Since original work is very hard to find and was published only in Russian language, we will present the short description of the method.
